### 2022-05-16 readme.Rmd
update to readme - remove deldir bit as we don't use it anymore.

---
output:
  md_document:
    variant: gfm
---

[![Travis build status](https://travis-ci.org/skiptoniam/ppmData.svg?branch=master)](https://travis-ci.org/skiptoniam/ppmData)

## ppmData is an R package that can be used to set up a quadrature scheme for multiple species point processes modelling and could be used in [ecomix](https://github.com/skiptoniam/ecomix) or for other multiple species Point Processes.

The approach uses quasi-random sampling to generate a quadrature
scheme based (e.g Berman & Turner 1992; Foster et al, 2018). Quasi-random 
sampling quadrature are form of spatially-balanced survey design or 
point stratification that aims to reduce the frequency of placing samples 
close to each other (relative to pseudo-random or grid designs). A quasi-random 
quadrature design improves efficiency of background point sampling 
(and subsequent modelling) by reducing the amount of spatial 
auto-correlation between data implying that each sample is providing as much 
unique information as possible (Grafston & Tille, 2013, Foster et al., 2018)
and thus reducing low errors for geostatistical prediction (Diggle & Ribeiro, 2007).
Because the quasi-random design is not on a regular grid we use a Dirichlet 
tessellation to generate polygons for each point. Areal weights are then derived
from these polygons.

### References

Diggle, P. J., P. J. Ribeiro, Model-based Geostatistics. Springer Series in Statistics. Springer, 2007.

Foster, S.D., Monk, J., Lawrence, E., Hayes, K.R., Hosack, G.R. and Przeslawski, R., 2018. Statistical Considerations for Monitoring and Sampling.

Grafstrom, Anton, and Yves Tille. "Doubly balanced spatial sampling with spreading and restitution of auxiliary totals." Environmetrics 24.2 (2013): 120-131.

Warton, D. I., and L. C. Shepherd. "Poisson point process models solve the pseudo-absence problem for presence-only data in ecology." The Annals of Applied Statistics 4.3 (2010): 1383-1402.