### 2022-05-16 readme.Rmd
I HAVE BROKEN THIS!

---
output:
  md_document:
    variant: gfm
---

[![Travis build status](https://travis-ci.org/skiptoniam/ppmData.svg?branch=master)](https://travis-ci.org/skiptoniam/ppmData)

## ppmData is an R package that can be used to set up a quadrature scheme for 
multiple species point processes modelling and could be used in 
[ecomix](https://github.com/skiptoniam/ecomix) or for other multiple species 
Point Processes.

The approach uses quasi-random sampling (Grafston & Tille, 2013; Foster et
al, 2018) to generate a quadrature scheme for presence-only modelling 
(Berman & Turner 1992). Quasi-random sampling quadrature are form of spatially-
balanced survey design or point stratification that aims to reduce the frequency
of placing samples close to each other (relative to pseudo-random or grid 
designs). A quasi-random quadrature design improves efficiency of background 
point sampling (and subsequent modelling) by reducing the amount of spatial 
auto-correlation between data implying that each sample is providing as much 
unique information as possible (Grafston & Tille, 2013, Foster et al., 2018)
and thus reducing low errors for geostatistical prediction (Diggle & Ribeiro,
2007). Because the quasi-random design is not on a regular grid we use a 
Dirichlet tessellation to generate polygons for each point. Areal weights are 
then derived from these polygons.

### References

Diggle, P. J., P. J. Ribeiro, Model-based Geostatistics. Springer Series in 
Statistics. Springer, 2007.

Foster, S.D., Monk, J., Lawrence, E., Hayes, K.R., Hosack, G.R. and Przeslawski,
R., 2018. Statistical Considerations for Monitoring and Sampling.

Grafstrom, Anton, and Yves Tille. "Doubly balanced spatial sampling with 
spreading and restitution of auxiliary totals." Environmetrics 24.2 (2013): 120-131.

Warton, D. I., and L. C. Shepherd. "Poisson point process models solve the 
pseudo-absence problem for presence-only data in ecology." The Annals of Applied
Statistics 4.3 (2010): 1383-1402.